---
title: "PS811 Portfolio"
author: "Valeriia Umanets"
output:
  html_document: 
    toc: true ## table of contents
    code_folding: hide
---
```{r setup, include=FALSE}
## Defaults for R chunks
## Add any R packages 

knitr::opts_chunk$set(echo = FALSE, ## echo = TRUE means code will show
                      warning=FALSE, ## supress warnings and messages from R
                      message=FALSE,
                      fig.path='Figs/', ## where to save figures
                      fig.height = 3, ## default figure size (can differ in each chunk)
                      fig.width = 7)

## Rrequired packages. 
requires <- c("tidyverse",
              "magrittr",
              "foreign",
              "readstata13",
              "here",
              "ggplot2",
              "ggpubr",
              "dplyr",
              "bindrcpp")

## Install all other from CRAN
to_install <- c(requires %in% rownames(installed.packages()) == FALSE)
install.packages(c(requires[to_install], "NA"), repos = "https://cloud.r-project.org/" )



## Load all required R packages
library(ggplot2); theme_set(theme_minimal())
library(magrittr)
library(here)
library(gplots)
library(ggpubr)
library(tidyverse)
library(dplyr)
library(bindrcpp)

options(stringsAsFactors = F) # set this as the default
```

<!-- The above header sets everything up. -->





### Research Interests
I am interested in elections, parties, and women’s representation and political participation in non-democratic political regimes.

---

### Questions I hope to answer
What is the role of type of electoral system and form of governance is the success of hegemonic parties in electoral authoritarian regimes? Is women’s political participation related to the Number of Effective Parliamentary Parties? Are women more or less likely to succeed in regimes with a hegemonic party?

---

### Methods
I am interested in various regression models, particularly in panel data, cluster analysis, technics applied in natural experiments, text analysis, and qualitative methods such as interviews and focus groups.

---

### Relevant data
- The Authoritarian Regimes Dataset (ARD) is a comprehensive dataset on authoritarian regimes in the world between 1972-2014. It distinguishes different forms of authoritarian regimes by their institutional setups. ([data](https://sites.google.com/site/authoritarianregimedataset/home)) 
- The Database of Political Institutions (DPI) presents institutional and electoral results data such as measures of checks and balances, tenure and stability of the government, identification of party affiliation and ideology, and fragmentation of opposition and government parties in the legislature. ([data](https://publications.iadb.org/handle/11319/7408))
- V-Dem database is data with over 350 indicators on democratic and authoritarian systems across the world from 1900 to the present ([data](https://www.v-dem.net))

<!-- 
Valeriia,
These datasets look like solid options. They all seem quite large, so your key task will be to select a few variables of interests. 
Try writing a few lines of R code that loads these data from their URL and then uses the select() from dplyr to trim it down to the columns you want. Make sure to include any ID columns. 
If the files are very large, you may want to make a short .R script that loads from the URL, selects the variables you want, and then saves your new, smaller data file in a folder called "data" (make sure to save this script in your 811 folder, commit it, and push it). This portfolio page can then load this smaller data file from that folder using here(data/YourNewDataFile).
Best,
Devin
-->

```{r data, include=FALSE}
# please keep all data in a data subfolder
parties <- read.csv("data/umanetsdata.csv", stringsAsFactors = F)

d <- parties # shorter name = easier

# here is another way of organizing the data. Instead of just by country, splitting it out by source. How you structure these data depend on what you want to do with them, but your current figures compare sources, so it is helpful to have a source variable.
d %<>% 
  gather(enp_lt, enp_mol, enp_gol, # gather these variables into new name-value pairs
         key = "ENPsource", # the names/source of the enp scores
         value = "ENP") # the enp scores (values)

d$ENP %<>% as.numeric() # you now have one numeric vector with all the ENP scores 

## Adding this information to the data will mean you don't need to keep typing it. Also, makes it more clear for anyone looking at your data
d$ENPsource %<>% str_replace("enp_lt", "ENP by Laakso and Taagepera (1979)")
d$ENPsource %<>% str_replace("enp_mol", "ENP by Molinar (1991)")
d$ENPsource %<>% str_replace("enp_gol", "ENP by Golosov (2010)")


glimpse(d)
```


### Descriptive statistics of three measurements of the ENP.


####Fig 1:  Density Plots for Effective Number of Parties.
```{r, echo= TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
require(gridExtra)

density1 <- parties %>% 
	  ggplot() + 
	  aes(x = enp_lt) +
	  geom_density() + 
	  labs(x = "ENP by Laakso and Taagepera (1979)")
density2 <- parties %>% 
	  ggplot() + 
		aes(x = enp_mol) +
	  geom_density() + 
	  labs(x = "ENP by Molinar (1991)")
density3 <- parties %>% 
	  ggplot() + 
	  aes(x = enp_gol) +
	  geom_density() + 
	  labs(x = "ENP by Golosov (2010)")
grid.arrange(density1, density2, density3, ncol=3,
             top = (""))
```


####Fig 2: Scatter Plots for Effective Number of Parties.
```{r, echo= TRUE}
library(dplyr)
library(ggplot2)
require(gridExtra)

scatter1 <- parties %>% 
  ggplot() + 
  aes(x = year, y = enp_lt) + 
  geom_point(alpha = .2) +
  labs(x = "Year", y = "ENP by Laakso and Taagepera (1979)")
scatter2 <- parties %>% 
  ggplot() + 
  aes(x = year, y = enp_mol) + 
  geom_point(alpha = .2) +
  labs(x = "Year", y = "ENP by Molinar (1991)")
scatter3 <-  parties %>% 
  ggplot() + 
  aes(x = year, y = enp_gol) + 
  geom_point(alpha = .2) +
  labs(x = "Year", y = "ENP by Golosov (2010)")
grid.arrange(scatter1, scatter2, scatter3, ncol=3,
             top = (""))

```

####Fig 3: Scatter plot and boxplots of ENP by Laakso and Taagepera by type of electoral system. 
```{r, echo= TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

parties %>%
  mutate(elecsys = ifelse(elecsys==1,'Proportional Representation',elecsys)) %>%
  mutate(elecsys = ifelse(elecsys==0,'Mixed Electoral System',elecsys)) %>%
  mutate(elecsys = ifelse(elecsys==2,'Plurality',elecsys)) %>%
  filter(elecsys!=".") %>% 
  ggplot() + 
  aes(x = elecsys, y = enp_lt, colour=elecsys)+
  geom_boxplot(colour="black")+
  geom_jitter(height = 0, alpha = .5)+
  scale_colour_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))+
  labs(x= "", y = "ENP by Laasko and Taagapera", colour = "Electoral systems")+
  ggtitle("") 

  
```


###Other Plots



``` {r, echo = T, results = 'hide'}
par(mfrow=c(1,3))
hist (parties$enp_lt, freq = FALSE, breaks = 20, col = "grey",
      main = "",
      xlab = "ENP by Laakso and Taagepera (1979)")
hist (parties$enp_mol, freq = FALSE, breaks = 20, col = "grey",
      xlab = "ENP by Molinar (1991)",
      main = "")
hist (parties$enp_gol, freq = FALSE, breaks = 20, col = "grey",
      xlab = "Golosov (2010)",
      main = "")
mtext("Histograms for Effective Number of Parties", side = 3, line = -2, outer = TRUE)
``` 

```{r density}
# Here is density with regular ggplot
d %>% 
  ggplot() + 
  aes(x = ENP) + 
  geom_density() + 
  facet_wrap("ENPsource") + # facet_grid and facet_wrap allow you to break out your data by groups in one plot
  labs(x = "Effective Number of Parties")
```

```{r histogram}
# Here is histograms with regular ggplot
d %>% 
  ggplot() + 
  aes(x = ENP) + 
  geom_histogram() + 
  facet_wrap("ENPsource") + 
  labs(x = "Effective Number of Parties",
       y = "Number of Countries")

```


```{r, echo = T, results = 'hide'}
par(mfrow = c(1,3))
plot(parties$year, parties$enp_lt, pch = 16,
     xlab = "Year",
     ylab = "ENP by Laakso and Taagepera (1979)")
plot(parties$year, parties$enp_mol, pch = 16,
     xlab = "Year",
     ylab = "ENP by Molinar (1991)")
plot(parties$year, parties$enp_gol, pch = 16,
     xlab = "Year",
     ylab = "ENP by Golosov (2010)")
mtext("Scatter plots for Effective Number of Parties", side = 3, line = -2, outer = TRUE)
```

```{r scatter}
# Here is scatter plot with regular ggplot
d %>% 
  ggplot() + 
  aes(x = year, y = ENP, color = ENPsource) + 
  geom_point(alpha = .2) +
  geom_smooth(method = "lm") + 
  labs(x = "Effective Number of Parties")
```

```{r average-enp}
# Here is scatter plot with regular ggplot
d %>% 
  group_by(year, ENPsource) %>% # by piping in d, rather than using ggplot(d), we can modify the data before plotting without creating a new object. Here, we get the mean ENP: 
  summarise(Mean = mean(ENP)) %>% 
  ggplot() + 
  aes(x = year, y = Mean, color = ENPsource) +
  geom_line() +
  labs(x = "Effective Number of Parties")
```


```{r, echo = T, results = 'hide'}
par(mfrow=c(1,3))
boxplot(parties$enp_lt, ylab = "ENP by Laakso and Taagepera (1979)")
boxplot(parties$enp_mol, ylab = "ENP by Molinar (1991)")
boxplot(parties$enp_mol, ylab = "ENP by Golosov (2010)")
mtext("Box plots for Effective number of parties", side = 3, line = -2, outer = TRUE)
```



